Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-12-16 demystify.tex
updates

\part{Your Application Materials}\label{part:application}

@@

\section{Publications} The most concrete evidence of research ability is having \emph{papers in reputable international journals or conferences}.
Having published papers, especially at top venues, is a sign that you have been successfully involved in research.

@@

In your SOP, focus on research potential and convince us through your experience, e.g., published papers. Back up your claims with \textbf{concrete evidence}. For example, if you say you have experience with teaching, then show what you did, e.g., undergrad TA or mentoring someone.  If you say you work on a project, then show some results, e.g., paper submitted (or even rejected), achieved certain performance improvement over the state of the art.

@@

Finally, your SOP reviewed by your LoR writers and professors, especially those who have served in adcom, or even postdocs or PhD students as they have been through this process.

@@

\begin{commentbox}[Vu:]
One common pitfall is to write about how you are from a poor family, a small town, or a developing country.  These are not related to DEI and raise concerns about you not understanding or even bother to do some basic research about the topic.   I have seen people consider themselves as minority because they are Asian or international students studying in the US (have you seen how many Asian or international students are in CS programs in the US?).\\

Moreover, do not make up stories or exaggerate your experiences.  Just show your awareness and understanding of the topic and what you did or how you plan to contribute to it (e.g., mentoring students, helping others). One of the best DEI statements I have read starts the acknowledgement that the applicant is not a minority, but a white male in a STEM field (i.e., that make them the majority), and that they do not have much experience in DEI, but show that they are aware of the challenges faced by minority groups and how they can contribute to it.
\end{commentbox}

>>>>

\part{Application Materials}\label{part:application}

@@

\section{Publications}\label{sec:publications}

The most concrete evidence of research ability is having \emph{papers in reputable international journals or conferences}.
Having published papers, especially at top venues, is a sign that you have been successfully involved in research.

@@

In your SOP, focus on research potential (\autoref{chap:research-experience}) and convince us through your experience, e.g., published papers (\autoref{sec:publications}). Back up your claims with \textbf{concrete evidence}. For example, if you say you have experience with teaching, then show what you did, e.g., undergrad TA or mentoring someone.  If you say you work on a research project, then show some results, e.g., paper submitted (or even rejected), achieved certain performance improvement over the state of the art.

@@

Finally, have your SOP reviewed by your LoR writers (\autoref{sec:help-your-LOR-writers}) and professors, especially those who have served in adcom, or even postdocs or PhD students as they have been through this process.

@@

\begin{commentbox}[Vu:]
One common pitfall is to write about how you are from a poor family, a small town, or a developing country.  These are not related to DEI and show that you do not understand the topic.   I have seen people consider themselves as minority because they are Asian or international students studying in the US (have you seen how many Asian or international students are in CS programs?).\\

Moreover, do not make up stories or exaggerate your experiences.  Just show your awareness and understanding of the topic and what you did or how you plan to contribute to it (e.g., mentoring students, helping others). One of the best DEI statements I have read starts the acknowledgement that the applicant is not a minority, but a white male in a STEM field (i.e., they are the majority), and that they do not have much experience in DEI, but show that they are aware of the challenges faced by minority groups and how they can contribute to it.
\end{commentbox}
